---
title: "Stat 243 -- Homework 05"
author: "Richmond Yevudza"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  pdf_document:
    fig_height: 2.2
    fig_width: 4
  html_document:
    fig_height: 2.2
    fig_width: 4
  word_document:
    fig_height: 2.2
    fig_width: 4
---

```{r, setup, include = FALSE, message=FALSE}
# load packages that are going to be used
require(mosaic)      # this loads ggformula (for plotting), etc. too
require(Lock5withR)  # this loads data sets from our book

# Some customization. You can alter or delete as desired (if you know what you are doing).

theme_set(theme_bw())     # change theme for ggplot2/ggformula

knitr::opts_chunk$set(
  tidy = FALSE,     # display code as typed (rather than reformatted)
  size = "small")   # slightly smaller font for code
```

<!-- Put your work below here.  Put text in text chunks, code in R chunks. -->

## 8.1

It can be seen that variation of the values in group 1 and group 2 in a dataset A is large than dataset B. So the difference in the group of dataset is uncertain. But there is very less variation in dataset B. So, it is certain that groups are from a dissimilar population. Hence, the sample data B has a stronger evidence of a difference in the two population means.

## 8.8

**Four groups with $n_1$ = 10, $n_2$ = 10, $n_3$ = 10, and $n_4$ = 10. ANOVA table is as follows:**

```{r 8point8question, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
problemTable88 <- "
| Source  | df | SS   | MS   | F-statistic |
|---------|----|:----:|------|:-----------:|
| Groups  | 3  | 960  | 320  | 2           |
| Error   | 36 | 5760 | 160  |
| Total   | 39 | 6720 |
"
cat(problemTable88) # output the table in a format good for HTML/PDF/docx conversion
```

## 8.10

**Three groups with $n_1$ = 5, $n_2$ = 8, $n_3$ = 7, and $n_4$ = 5. ANOVA table is as follows:**

```{r 8point10question, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
problemTable810 <- "
| Source  | df | SS   | MS      | F-statistic |
|---------|----|:----:|---------|:-----------:|
| Groups  | 3  | 600  | 200     | 5.25        |
| Error   | 21 | 800  | 38.0952 |
| Total   | 24 | 1400 |
"
cat(problemTable810) # output the table in a format good for HTML/PDF/docx conversion
```

## 8.12a

5

## 8.12b

NULL = means of all groups are equal

ALTERNATIVE = means of at least two groups are different

## 8.12c

0.0012

## 8.12d

we reject $H_0$ due to p-value being less than level of significance, so there's enough evidence to conclude at least two group means are different

## 8.14a

4

## 8.14b

NULL = means of all groups are equal

ALTERNATIVE = means of at least two groups are different

## 8.14c

0.538

## 8.14d

we fail to reject $H_0$ due to p-value being greater than level of significance, so there's not enough evidence to conclude at least two group means are different

### 8.16a

People perform differently on achievement tasks based on exposure to different colors.

### 8.16b

```{r 8point16question, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
problemTable816 <- "
| Source  | df | SS    | MS      | F-statistic |
|---------|----|:-----:|---------|:-----------:|
| Groups  | 2  | 27.7  | 13.85   | 16.49       |
| Error   | 68 | 57    | 0.84    |
| Total   | 70 | 84.70 |
"
cat(problemTable816) # output the table in a format good for HTML/PDF/docx conversion
```

### 8.16c

0

### 8.16d

There is evidence showing that color affects achievement performance, as the p-value was lower than the level of significance.

## 8.24a

bright light

## 8.24b

Yes

## 8.24c

Yes

## 8.24d

The variables are LL, LD, DM, and BM gain

LL, LD, and LL are all categorical

BM gain is a quantitative variable

## 8.25a

$H_0$ = amount of light at night doesn't affect weight gain

$H_1$ = amount of light at night has some affect on weight gain

## 8.25b

F: 8.38
P: 0.0002
Conclusion: reject null hypothesis $H_0$ due to p-value being lower than level of significance

## 8.25c

Yes - the higher the light level, the more weight the mice gained. LL had the highest weight gain as a result.

## 8.25d

Yes, since these are results from a randomized experiment.

## 8.26

$H_0$ = activity level mean insignificant

$H_a$ = activity level mean significant

F: 0.09
P: 0.910

Conclusion: P greater than 0.1, so accept $H_0$, therefore there's no significant difference in activity level.

## 8.27a

Nope, standard deviations are too different

## 8.27b

Don't reject $H_0$.

## 8.31

Conditions for ANOVA make sense

F = 27.86

P = 0.000

Reject $H_0$

## 8.46a

No difference

## 8.46b

No difference 

## 8.46c

There is a difference; the null hypothesis is being rejected

## 8.52

```{r 8point52question, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
problemTable852 <- "
| Anova SF
|---------|-------|-------|---------|----------|
| Groups  | Count | Sum   | Average | Variance |
| Low     | 30    | 2055  | 68.5    | 263.569
| Medium  | 30    | 1760  | 58.6667 | 204.023
| High    | 30    | 1745  | 58.1667 | 189.7989
|---------|-------|-------|---------|----------|
|---------|-------|-------|---------|----------|
| ANOVA
|---------------------|----|----|----------|----------|
| Source of Variation | SS | df | MS       | F        | p-value  | F-crit
| Between Groups      | 30 | 2  | 1018.611 | 4.648427 | 0.012077 | 3.101296
| Within Groups       | 30 | 87 | 219.1303
| Total               | 30 | 89
"
cat(problemTable852) # output the table in a format good for HTML/PDF/docx conversion
```

since p-value is greater than 0.05, means are in fact significant

significance level: 0.05
number of treatments: 3
DF error: 87
MSE: 219.130
q-stat value: 3.4000

```{r 8point52againquestion, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
problemTable852again <- "
| Anova SF
|---------------|----------|-------------|-------------|--------|
| pop mean diff | crit val | lower limit | upper limit | result |
| mu1-mu2       | 9.83     | 9.19        | 0.64        | 19.02  | means are different
| mu1-mu3       | 9.83     | 9.19        | 0.64        | 19.02  | means are different
| mu2-mu3       | 0.00     | 9.19        | -9.19       | 9.19   | means are NOT different
"
cat(problemTable852again) # output the table in a format good for HTML/PDF/docx conversion
```