---
title: "Stat 243 -- GS01"
author: "Richmond Yevudza"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  pdf_document:
    fig_height: 2.2
    fig_width: 4
  html_document:
    fig_height: 2.2
    fig_width: 4
  word_document:
    fig_height: 2.2
    fig_width: 4
---

```{r, setup, include = FALSE, message=FALSE}
# load packages that are going to be used
require(mosaic)      # this loads ggformula (for plotting), etc. too
require(Lock5withR)  # this loads data sets from our book

# Some customization. You can alter or delete as desired (if you know what you are doing).

theme_set(theme_bw())     # change theme for ggplot2/ggformula

knitr::opts_chunk$set(
  tidy = FALSE,     # display code as typed (rather than reformatted)
  size = "small")   # slightly smaller font for code
```

<!-- Put your work below here.  Put text in text chunks, code in R chunks. -->

#### 1.12a 
```{r}
countries <- data.frame(
    Variable = c("Country", "Code", "LandArea", "Population", "Energy", "Rural",
                 "Military", "Health", "HIV", "Internet", "Developed", "BirthRate",
                 "ElderlyPop", "LifeExpectancy", "CO2", "GDP", "Cell", "Electricity"),
    VariableType = c("Categorical", "Categorical", "Quantitative", "Quantitative",
                     "Quantitative", "Quantitative", "Quantitative", "Quantitative",
                     "Quantitative", "Quantitative", "Categorical", "Quantitative",
                     "Quantitative", "Quantitative", "Quantitative", "Quantitative",
                     "Categorical", "Quantitative")
)
countries
```

#### 1.12b
1. What is the average population of all countries around the world?
2. Which country has the highest GDP?

#### 1.12c
1. Do the countries with larger area have more access to the internet?
2. Do the countries with smaller area have a lower birth rate?

#### 1.20a
The given dataset is about the number of days to row alone the Atlantic Ocean. The number of cases are 8.
There are two variables in the given study, namely gender and number of days to cross the Atlantic Ocean.
For the Gender variable, the subjects are classified into two groups (male and female). Therefore it is categorical.
For the number of days to cross the Atlantic Ocean, the data represents a numerical value. Therefore, it is quantitative.

#### 1.20b
```{r}
Atlantic <- data.frame(
    Gender = c("M", "M", "M", "M", "M", "F", "F", "F"),
    DaysToRow = c(40, 87, 78, 106, 67, 70, 153, 81)
)
Atlantic
```

#### 1.22a
The information is obtained from 41 people who participated in the study. Therefore there are 41 cases in this study.

#### 1.22b
The variables are as following:

* Whether or not the participants were involved in mediation program (categorical)
* Brain wave activity across the front of the left hemisphere measured before study (quantitative)
* Brain wave activity across the front of the left hemisphere measured after study (quantitative)
* Brain wave activity across the front of the left hemisphere measured 4 months later (quantitative)
* Immune response to the vaccine after 1 month (quantitative)
* Immune response to the vaccine after 2 months (quantitative)
* Positive emotions before (quantitative)
* Negative emotions before (quantitative)
* Positive emotions after (quantitative)
* Negative emotions after (quantitative)

#### 1.22c
The explanatory variable is a factor that has been manipulated in the experiment by the researcher. In this study, **whether or not the participants were involved in mediation program** is the explanatory variable.

#### 1.22d
If it is assumed that each case represents a row, then the dataset will contain **41 rows**. If each variable is a column, then the dataset will contain **at least 11 columns**.

#### 1.24
Data might be collected from a sample of people who are eligible to vote and the following procedure could be followed:

* Ask about the person's political party.
* Ask whether they voted in the last election.

In this dataset, the cases are people who are eligible to vote. The variables are the person's political party, and whether or not they voted in the last election.

#### 1.26
**Question:** To what extent does your college GPA affect your work salary in the future?

To be able to come up with data for this question I will have to gather a sample of earners from different earning classes (i.e 60-99k, 100-199k, 200-299k...) and find out their GPA. By comparing these two quantitative variables, I might be able to come up with a conclusion.

The cases in this dataset will be the workers. The variables will be GPA and salary.

#### 1.50
This method of data collection is **not biased**.

This is mainly because in this sampling method, a random sample of one type of typewriter is selected and testes for the number of pages that can be printed before the ink runs out. Then after that, teh average number of pages for that can be printed before the ink turns out is estimated for that selected type of typewriter. It is quite clear that the selection is random and the results that are gotten can also be trusted.

#### 1.52
This statistic can be generalized to **all parents in Kansas City**.

#### 1.56
The sample is **not the representative** of all Australians.

This is mainly because the sample is a volunteer sample and therefore the participants are not a representative of the population. In addition, the advertisements for the study were on two rock radio stations in Sydney. This means that it is only the people who hear about it that have the chance to participate and clearly not the whole population hear about it. Also, there are participants who choose because they like alcohol and marijuana. These people tend to know more about the substances than the overall population.

#### 1.58a
300 tanning salons.

#### 1.58b
**Yes**, the sample is representative of all tanning salons in the US because the sample is selected in such a way that shows this.

#### 1.58c
Although the sample is random, the results don't give an accurate picture of the dangers of tanning because most of the salons are concerned more about marketing what they offer than actually showing facts. So quite a number of the responses given by the salons are not true.

#### 1.58d
**Yes**, the study accurately portrays the messages tanning salons give to teenage girls because the sample is selected well.

#### 1.60a
All US residents.

#### 1.60b
All the emergency room patients in the US.

#### 1.60c
i. NHANES
ii. NHAMCS
iii. NHAMCS
iv. NHANES

#### 2.13a
**Sample**: The random 119 players who were observed playing Rock-Paper-Scissors.

**Population**: All people who play Rock-Paper-Scissors.

The variable measures the option selected (rock, paper or scissors) which is categorical data. It also measures the frequency which is quantitative data.

#### 2.13b
```{r}
RockPaperScissors <- data.frame(
    OptionSelected = c("Rock", "Paper", "Scisors"),
    Frequency = c(66, 39, 14)
)
data.frame(
  OptionSelected = RockPaperScissors$OptionSelected, 
  Frequency = RockPaperScissors$Frequency,
  RelativeFrequency = RockPaperScissors$Frequency/sum(RockPaperScissors$Frequency)
  )
```

#### 2.13c
For the odds to be in your favor, your best bet would be to play **paper** as rock's relative frequency is 0.55 (the highest of them all).

#### 2.13d
The option picked for the next round should be **scissors** because the players are more likely to repeat the option paper.

#### 2.18a
```{r}
TwoWayTable <- matrix(c(363, 176, 196, 735, 557, 466, 789, 1812, 20, 26, 32, 78, 940,
                        668, 1017, 2625), ncol=4, byrow=TRUE)
rownames(TwoWayTable) <- c("Agree", "Disagree", "Don't know", "Total")
colnames(TwoWayTable) <- c("HS", "Some", "College", "Total")
TwoWayTable
```

#### 2.18b
Required percentage $=\frac{363}{940} = 0.386 = 38.6$% 

Therefore, the percentage of agree with high school degree or less is **38.6%**.

Required percentage $=\frac{176}{668} = 0.263 = 26.3$% 

Therefore, the percentage of agree with some college is **26.3%**.

Required percentage $=\frac{196}{1017} = 0.193 = 19.3$% 

Therefore, the percentage of agree with college graduate or higher is **19.3%**.

#### 2.18c
Required percentage $=\frac{1017}{2625} = 0.387 = 38.7$% 

Therefore, **38.7%** of the people participating in the survey have a college degree or higher.

#### 2.18d
Required percentage $=\frac{557}{1812} = 0.307 = 30.7$% 

Therefore, **30.7%**  of people who disagreee with the statement have a college degree or higher.

#### 2.19a
Percentage of female smokers in female sample,

$\frac{16}{169} = 0.095 = 9.5$% 

Percentage of male smokers in male sample,

$\frac{27}{193} = 0.140 = 14.0$% 

Therefore, the gender with the higher percentage of smokers is the **male**.

#### 2.19b
Required proportion $=\frac{43}{362} = 0.119 = 11.9$% 

Therefore, the proportion of smokers for the whole sample is **11.9%**.

#### 2.19c
Required proportion $=\frac{16}{43} = 0.372 = 37.2$% 

Therefore, the proportion of smokers in the sample who are female is **37.2%**.

#### 2.20a
This experiment is an observational study. The two variables are.**Does the person have dyslexia or not** and **Does the person have the DYXCI break or not**.

#### 2.20b
The total number of "Dyslexia group" is 109 and the total nuber of "Control group" is 195. Therefore, the total number of rows (cases) in the study is **304**.

The total number of columns is**2** as there are two variables in this data set.

#### 2.20c
```{r}
TwoWayTable <- matrix(c(10, 99, 109, 5, 190, 195, 15, 289, 304), ncol=3, byrow=TRUE)
rownames(TwoWayTable) <- c("Dyslexia group", "Control group", "Total")
colnames(TwoWayTable) <- c("Gene break", "No break", "Total")
TwoWayTable
```

#### 2.20d
Proportion of "Dyslexia group" with "Gene break" $=\frac{10}{109} = 0.902$ 

Proportion of "Control group" with "Gene break" $=\frac{5}{195} = 0.026$ 

Therefore, looking at the proportion of each group who have the break on DYXCI gene, a conclusion can be drawn that there appears to be a very substantial difference between the group with dyslexia and the control group.

#### 2.20e
There appears to be an association between the genetic marker and dyslexia for the people in this sample.

#### 2.20f
If the association appears to be strong, then it can't be assumed that the gene distribution causes dyslexia because the data set comes from the observational study but not from an experiment.

#### 2.55a
The shape of the distribution of ratings of word is **left skewed** because the left side of the figure is extended larger than the right side.

#### 2.55b
Looking at Figure 2.14, the value that separates the area into almost two equal parts is 6.5. Therefore, the median value is around **6.5**.

#### 2.55c
The value of the mean will be **smaller** than the median because the shape of the shape of the distribution of ratings of word is skewed to the left which shows that the mean is smaller than the median.

#### 2.60a
$\overline x_f = \frac{\sum(x)}{n} = \frac{64}{10} = 6.4$

Therefore, the mean number of hours spent exercising by the females is **6.4**.

#### 2.60b
$\overline x_m = \frac{\sum(x)}{n} = \frac{177}{26} = 6.81$

Therefore, the mean number of hours spent exercising by the males is **6.81**.

#### 2.60c
$\overline x_m - \overline x_f = 6.81 - 6.4 = 0.41$ 

This shows that the average hours spent on exercise by the males is higher than the average hours spent on exercise by the females by **0.41**.

#### 2.63a
Data set: 1, 49, 50, 52, 56

$\overline x = \frac{\sum(x)}{n} = \frac{208}{5} = 41.6$

Median: 50

Mean: 41.6

#### 2.63b
Data set: 1, 20, 30, 80, 100

$\overline x = \frac{\sum(x)}{n} = \frac{231}{5} = 46.2$

Median: 30

Mean: 46.2

#### 2.82
* Minimum: 58
* First quartile: 65
* Median: 68
* Third quartile: 70
* Maximum: 77

#### 2.91
The required interval is $\overline x \pm 2s$, where $\overline x$ is the mean and s is the standard deviation.

$\overline x \pm 2s = 200 \pm 2(25) = 200 \pm 50 = (150, 250)$

Therefore, the 95% of the data values lies between **150 and 250**.
















